---
title: "Assignment 3, part 3"
output: 
  html_document: 
    keep_md: yes
    df_print: paged
---

__Name:__ Zun Wang

__Student ID:__ 915019847



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, max.print=100)
```

**Exercise 22:**
Scroll through the alignment.  Are there any additional regions where you are suspicious of the alignment?  If so, list when and state the potential problem.
```{r}
#Around 1700bp, there is a gap rich region, where the alignment does not work well.
```

**Exercise 23:** 
```{r}
library(tidyverse)
library(Biostrings)
```

```{r}
inpath <- "../output/mafft_maxiter100_195_op.5.fa"
outpath <- "../output/mafft_maxiter100_195_op.5_trimmed_75pct.fa"
alignment <- readDNAMultipleAlignment(inpath)
alignment
```
```{r}
alignment <- DNAMultipleAlignment(alignment,start=1000,end=48250)
```
```{r}
alignment <- maskGaps(alignment, min.fraction=0.25, min.block.width=1)
maskedratio(alignment) 
alignment <- alignment %>% as("DNAStringSet")
newnames <- names(alignment) %>% 
  tibble(name=.) %>%
  mutate(name=str_replace_all(name," ","_")) %>% #replace " " with "_" because some programs truncate name at " "
  separate(name, 
           into=c("acc", "isolate", "complete", "name", "country", "host"),
           sep="_?\\|_?") %>%
  mutate(name=str_replace(name, "Middle_East_respiratory_syndrome-related","MERS"),   # abbreviate
         name=str_replace(name, "Severe_acute_respiratory_syndrome-related", "SARS"), # abbreviate
         newname=paste(acc,name,country,host,sep="|")) %>% # select columns for newname
  pull(newname) #return newname
```
```{r}
head(newnames)
names(alignment) <- newnames
alignment %>% writeXStringSet(outpath)
```
__A:__ What is the sister taxon to Seq_H? What is the host for the virus in this group (provide the Latin and common names)
_hint: if you are having trouble finding Seq H in the tree, search for it using the (Aa) magnifying glass_
```{r}
#It is the MG772933 sequence. Its host is Rhinolophus Sinicus, which is the Chinese rufous horseshoe bat.
```
__B:__ Consider Seq_H plus its sister taxon as defining one taxonomic group of two species.  Look at the sister taxa to this group.  What is a general description for the viruses in this group?  List at least 3 hosts found in this group.
```{r}
#They are either SARS coronavirus or bat coronavirus. The hosts include Rhinolophus ferrumequinum, Rhinolophus affinis, and Chiroptera.
```

__C:__ Examine the sister group to the taxa described in parts A and B.  What are its host(s)?
```{r}
#Rhinolophus blasii.
```
__D:__ What do you think the host of the most recent common ancestor to the viruses in parts A and B was?
```{r}
#It is the Rhinolophus sinicus.
```
__E:__ Do you think that Seq_H evolved from a virus with a human host?  Why or why not?  If not, what did it evolve from?
```{r}
#It is most likely that this virus arise from bat instead of human host, since its related sequences are all from bats.
```